---
title: "What is the Recent Trend of Soccer in the Top Soccer Leagues"
author: "Eung Kyu Kim"
date: "December 20, 2020"
output: pdf_document
---
#Github Repository:
# https://github.com/williamkim1102/Recent-Trend-of-Soccer

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```


#Abstract 
  
  In this report, an analysis will be performed to find the recent trend of soccer in the German, English and Spanish Leagues from 2019-2020. Variables related to soccer will be used in linear regression models to find the relationship between the variables. From these models, a recommendation of how soccer should be played nowadays will be given. The data are from the official websites of the three leagues. 



#Keywords 

    + Soccer Trend 
    + Data Analysis 
    + Linear Regression 
    + Tactics
    + Multiple Linear Regression
    + How to earn points in soccer

#Introduction 

  Each country in the world has different opinions and political views, so the countries have been fighting against each other. Then what is the best thing that unites the world? One of the best ways to unite the world is through sports, especially, football. Football, as known as soccer in North America, is the most popular sport in the world and every country plays soccer. However, soccer is one of the sports that is extremely hard to predict the outcomes, which means there are lots of upsets and uncertainties. Recently, famous soccer teams' revenues have grown, so they have been spending more money to buy players (Connelly, 2020). However, soccer is not always about money and there are many other factors that can affect the outcomes. 

  The trends of soccer has been constantly changing. In the 1970s, a tactic called 'Total Football' was played by the Dutch soccer team. This tactic was all about adopting the roles of any other player in the team. (Siregar, 2018) This tactic required constantly running around the field, which means the players had to cover a lot of distance and it required stamina. After decades, around the early 2010s, a tactic called 'Tiki Taka' was the trend of soccer. Tiki Taka was used by the Spanish League team, FC Barcelona and the Spanish National Team and it was all about short passing and possession (Bairner, 2020). This tactic was one of the most influential tactic in soccer history and this made a lot of soccer teams to play with having high possession. 
  
  The most recent influential tactic of soccer was played by an English team called Leicester City FC in 2016. Their tactic also completely changed the way of thinking about soccer. Due to the influence of Tiki Taka, in the middle of 2010s, almost all the teams tried to have high possession and increase the number of short passing. On the other hand, Leicester City FC thought differently about how they should play football. Their tactic was about counter attacking, which means they were mostly defending for most of the times, but when they gain possession, they quickly passed the ball to the front and scored. Leceister City FC won the league that year, but compared to the other 19 teams in the league, their possession was 42.6%, which was the 3rd lowest in the entire league (Premier League, 2015-2016). This proved that possession does not define earning points (wins and draws). 
  
  As shown above, there have been different tactics being the trend in different years of time. Then, what is the current trend of football and what are the factors that make you earn points in the league? In this analysis, which factor affects the most for a football team to win points will be discussed through analyzing multiple data related to soccer. 

#Methodology 

##Data

  There were a total of three different data sets used in this analysis. Three of the data were sets of different leagues including the English Leauge, EPL, the German League, Bundesliga and the Spanish league, LaLiga. Each data set includes over 150 variables that are related to soccer. These data sets are data from the league of 2019-2020. Then, certain variables that are thought to be the most important factors to earn points were selected from the data sets. The variables that are chosen are 
  
  + Total_Pass 
  + Win 
  + Loss 
  + Draw
  + GF:Goals Scored
  + GA:Goals given
  + Possession
  + Expenditure
  
  Then, three leagues were combined into a one data set and it is shown in Table 1 in the Appendix ####. This combined data set will be used throughout the analysis and it will be modified in different sections. 

##Model

  When choosing the model of the analysis, two different regressions were used to find the best model. The first regression used is a simple linear regression and the other regression used is a mulitple linear regression. 

####Simple Linear Regression

  In a simple linear regression, a relationship between two variables can be found. The equation of a standard linear regression model is shown below.
$$
y = \beta_{0} + \beta_{1}*x_{1} + \epsilon
$$
The terms in the equation are: y, $\beta_{0}$, $\beta_{1}$, $x_{1}$ and e. 
Each representing:
+ y : Dependent Variable
+ $\beta_{0}$ : Intercept Term
+ $\beta_{1}$ : Slope Term
+ $x_{1}$ : Independent Variable
+ $\epsilon$ : Residual / Error term
  
These variables are found using the simple linear regression model and the values are used to find the relationship between the variables and the slope gives whether relation is positive or negative. Simple Linear Regression was used to find the relationship between these three different pairs of variables.

+ Points vs Expenditure
+ Points vs Total Number of Passes
+ Points vs Possession


####Multiple Linear Regression

  In a multiple linear regression, a relationship between a dependent variable and many independent variables is shown. When MLR is used, using the summary, you can find point estimates of the independent variables and using them, you can predict the outcome of a response variable. 
  
$$
y_{i} = \beta_{0} + \beta_{1}*x_{1} + \beta_{2}*x_{2} + ... + \beta_{i}*x_{i} + \epsilon
$$
The terms in the equation are: y, $\beta_{0}$, $\beta_{1}$, $x_{1}$ and e. 
Each representing:

+ $y_{i}$ : Dependent Variable
+ $\beta_{0}$ : Intercept Term
+ $\beta_{1}$, $\beta_{2}$, $\beta_{i}$  : Coefficients / Point Estimates
+ $x_{1}$, $x_{2}$, $x_{i}$ : Independent Variables
+ $\epsilon$ : Residual / Error term


#Results (3-4 Paragraphs)

##Simple Linear Regression Models

  The first model that was analyzed using the simple linear regression was finding the relationship between points earned and possession. As it has been mentioned earlier, one of the trends in the 2010s of soccer was having high percentage of possession, so the model was analyzed to check whether that is still correct for current trend of soccer. The Graph 1 in Appendix C shows the graph of Possession vs Points Earned. As it can be seen, there is a positive relationship between the two variables which means as the team gains more possession in games, they earn more points. The correlation of these two varaibles was found, which was 0.7435 and since the value is between -1 and 1 and not close to 0, it means they have high levels of associations between the two variables. Then, the coefficients of this model was found and using the coefficients in Appendix D, the estimated regression line can be written as

$$
 y = -35.594 + 1.722*x
$$
where x represents the possession and y represents the pointes earned. Using a simple linear regression, it has been found that points earned increase by 1.722 points as the team gains 1% of possession. 

  A model of finding the relationship between expenditure and points earned was analyzed using the simple linear regression. In graph 2 of the Appendix E, it shows a graph of the relationship between the expenditure of the teams and the points earned. Using the coefficients in Appendix F, we can conclude that the estimated regression line of the model is 

$$
 y = 41.5735 + 0.1232*x
$$  
where y is the points earned and x is the expenditure. However, it can be seen from Graph 2 that some teams' expenditure are significantly high compared to the other teams, so these values could affect the slope of the regression line to increase. Hence, another model was analyzed after filtering out 11 teams that had high expenditure. This model is shown in Graph 3 of Appendix G and using the coefficients in Appendix H, we can find that the estimated regression line of the model is 

$$
 y = 43.27156 + 0.07026*x
$$  
and we can see that the slope has decreased and from the graph it is visible that the slope is close to being horizontal. This informs us that after excluding teams who spend significantly high amount of money, the points eanred does not change by a lot when the teams spend more. Also, in the Appendix H, it shows that correlation between the two variable is about 0.18 and it is close 0, which means the two values are not correlated. Hence, the 'Expenditure_euro' variable will be excluded from the multiple linear regression that will be performed. 

##Multiple Linear Regression Models (MLR)

  After removing the expenditure variable from the dataset, mulitple linear regression was performed to different models. The variables that will be selected for multiple linear regression are

+ Shots
+ Total Number of Passes
+ Goals Given
+ Possession

and these variables were chosen because shots represent values for the attack, number of passes represent values for the midfielders, goals given are the values for the defenders and possession represents the entire team. Although there are many other variables for soccer, these variables were selected to represent the factors that can affect the gameplay. 

  The first MLR was done with using these four variables to find its effect on the pointes earned and this model can be seen in the Appendix I. By looking at the summary and the p values, we can see that the p-value of possession is significantly high (0.63731). Since this is greater than the alpha, 0.05, this means that the possession variable is not strong enough to suggest and effect exists in the points earned. Hence, the variable has to be excluded from the model.
  
  After removing the variable, MLR was used against to find the relationship between the variables and this model can be found in the Apeendix J. In this model, three variables: Passes, Shots, Goals Against were used and by looking at the p-values, it can be concluded that all of these three variables are highly related to the dependent variable, Point Earned. Additionally, the Multiple R-squared value was 0.8676 and this means that approximately 86.76% of variation in Points earned can be explained by our model. Lastly, using the coefficients of the model given in Appendix J, estimated regression line is found to be 
  
$$
 y_{Points Earned} = 54.7447 + 0.0011237 * x_{pass} + 0.0524 * x_{shots} -0.8505 * x_{Goals Given}
$$  

#Discussion 

##Summary

  Throughout this analysis, data of three soccer leagues: English, German and Spanish were used to find what factors affect the teams to earn points. These data were combined and certain variables that were thought to be relevent were chosen. Using this data two types of regression models were performed. In the simple linear regression models, expenditure variable and possession variable was used to find the relationship between the points earned and them. However, in the multiple linear regression model, total number of passes, shots taken and goals given were used to find the relationship between the points earned and them. 

##Conclusion

  In conclusion, the datasets of different soccer leagues were merged to increase the number of data and the accuracy of the analysis. Using this data, we found that possession and points earned have a positive relationship. Meanwhile it was found that expenditure doesn't affect the points earned by a lot after removing certain teams who spend billions of euros. This gave us a fact that current trend of soccer is not always about the money and the actual game play is more important than how much each team spends. 
  
  After doing the analysis of multiple linear regression with four variables, we have found that the possession actually does not affect how much the team earns points. From this, it can be concluded that recent trend of soccer has not changed since the 2016 which was the year that Leicester City FC won the league with having low percentage of possession. After removing the possession variable, it was found that the three variables: passes, shots, goals given are the most important factors for teams to earn points in soccer games. The regression line informs us that every pass the team makes, the points earned will increase by 0.00112 and every shot the team takes, the points earned will increase by 0.0524. On the other hand, the points will decrease by 0.8505 as the other teams score on the team. Also, from the Residuals vs Fitted Graph and the Normal Q-Q Graph in the Appendix K, it can be concluded that this multiple linear regression model gives a fairly good explanation of what factors affect the points earned. 
  
  From this analysis, it can tell the teams how to play soccer currently. The recent trend of soccer to earn points is, since the possession does not affect the results by much, try to pass the ball fastly and take a shot, but focus on defending. Again, from this, it can be proved that Leceister City FC's style of play is still the trend of soccer in the three leagues. Hence, the teams should focus on playing similar to Leceister City FC.

##Weakness & Next Steps

  Although linear regressions were used in this analysis, soccer is way more complicated than finding which factor affects the points earned. There could be attendance, team satisfactory, weather, wage and many other factors. However, only certain variables were chosen in models to come up with a conclusion, so that is the weakness of this analysis. For future analysis, more variables can be considered with more number of data sets, which can give an idea of how to run a football team with what kind of tactics to successfully lead a soccer team. 

#References

+ Bairner, R. (2020, May 8). What is tiki-taka? How tactics made famous by Barcelona and Spain work. What is tiki taka? How tactics made famous by Barcelona and Spain work | Goal.com. https://www.goal.com/en-ng/news/what-is-tiki-taka-barcelona-spain-tactics/5f3qumd4uank198jwik1ww8mr. 

+ Connelly, B. (2020, April 19). How soccer has changed in the past 10 years: From Mourinho's peak to reign of superclubs. ESPN. https://www.espn.com/soccer/english-premier-league/story/4086497/how-soccer-has-changed-in-the-past-10-years-from-mourinhos-peak-to-reign-of-super-clubs. 

+ Premier League. 2015-2016 Premier League Player Stats. FBref.com. https://fbref.com/en/comps/9/1467/stats/2015-2016-Premier-League-Stats. 

+ Siregar, C. (2018, December 14). What is Total Football? Famous tactics explained: the clubs, countries &amp; players to use it. https://www.goal.com/en/news/what-is-total-football-famous-tactics-explained-the-clubs/w5yd5dzofn4p17ewmav4atk2b. 


#Appendix 


##Appendix A

```{r, warning=FALSE, include = FALSE}
##Data Cleaning
library(tidyverse)

#Importing the general stats
EPL_19_20 <- read_csv("C:/Users/willi/Desktop/UT 20-21/STA304/Final/Data/EPL1920.csv")
Bundesliga_19_20 <- read_csv("C:/Users/willi/Desktop/UT 20-21/STA304/Final/Data/Bundesliga1920.csv")
LaLiga_19_20 <- read_csv("C:/Users/willi/Desktop/UT 20-21/STA304/Final/Data/LaLiga1920.csv")

#Importing other relevant datasets 
EPL_expenditure_19_20 <-read_csv("C:/Users/willi/Desktop/UT 20-21/STA304/Final/Data/EPLexpenditure1920.csv")
Bundesliga_expenditure_19_20 <-read_csv("C:/Users/willi/Desktop/UT 20-21/STA304/Final/Data/BundesligaExpenditure1920.csv")
LaLiga_expenditure_19_20 <-read_csv("C:/Users/willi/Desktop/UT 20-21/STA304/Final/Data/LaLigaExpenditure1920.csv")
```

```{r,warning=FALSE}

#Modifying the EPL expenditure numbers in the right way of writing because the units were in million
EPL_expenditure_Final <- EPL_expenditure_19_20 %>%
  mutate(Expenditure_euro = Expenditure * 1000000 ) %>% 
  mutate(Income_euro = Income * 1000000  ) %>%
  mutate(Balance_euro = Balance * 1000000 ) %>%
  select(Club, Expenditure_euro, Income_euro, Balance_euro, Arrivals, Departures)

#Modifying the Bundesliga expenditure numbers in the right way of writing because the units were in million
Bundesliga_expenditure_Final <- Bundesliga_expenditure_19_20 %>%
  mutate(Expenditure_euro = Expenditure * 1000000  ) %>% 
  mutate(Income_euro = Income * 1000000  ) %>%
  mutate(Balance_euro = Balance * 1000000 ) %>%
  select(Club, Expenditure_euro, Income_euro, Balance_euro, Arrivals, Departures)

#Modifying the LaLiga expenditure numbers in the right way of writing because the units were in million
LaLiga_expenditure_Final <- LaLiga_expenditure_19_20 %>%
  mutate(Expenditure_euro = Expenditure * 1000000  ) %>% 
  mutate(Income_euro = Income * 1000000  ) %>%
  mutate(Balance_euro = Balance * 1000000  ) %>%
  select(Club, Expenditure_euro, Income_euro, Balance_euro, Arrivals, Departures)
  

#Combining the expenditure dataset and general stats expenditure together
EPL1920 <- merge(EPL_19_20, EPL_expenditure_Final, by = "Club")
Bundesliga1920 <- merge(Bundesliga_19_20, Bundesliga_expenditure_Final, by = "Club")
LaLiga1920 <- merge(LaLiga_19_20, LaLiga_expenditure_Final, by = "Club")

```


##Appendix B

###Table 1

```{r}
##Combining the datasets together with the variables that will be used

EPL1920_tgt <- EPL1920 %>%
  select(Club, Total_Pass, Points, W, GF, GA, Shots, Poss, Expenditure_euro)

Bundesliga1920_tgt <- Bundesliga1920 %>%
  select(Club, Total_Pass, Points, W, GF, GA, Shots, Poss, Expenditure_euro)

LaLiga_tgt <- LaLiga1920 %>%
  select(Club, Total_Pass, Points, W, GF, GA, Shots, Poss, Expenditure_euro)

League_Combined_Prev <- rbind(EPL1920_tgt, Bundesliga1920_tgt)
League_Combined <- rbind(League_Combined_Prev, LaLiga_tgt)

kable(League_Combined)
```

##Appendix C

###Graph 1
```{r}
#Simple Linear Regression Comparing Points vs Total Possession


League_simple_poss <- ggplot(data = League_Combined, aes(x = Poss, y = Points )) +
  geom_point() + labs(title = "Relationship between Possession vs Points Earned",
                      y = "Points Earned",
                      x = "Possession")
League_simple_poss + geom_smooth(method = lm)  
 
```

##Appendix D
```{r}
#Relevant values to Graph 1, which are correlation and coefficients of Graph 1

coefficients_slr_poss <- lm(Points ~ Poss, data = League_Combined)
coefficients_slr_poss

cor(League_Combined$Poss, League_Combined$Points)


```

##Appendix E

###Graph 2
```{r}

#Simple Linear Regression Comparing Points vs Expenditure

League_Combined_expenditure <- League_Combined %>%
  mutate(expenditure_millions = Expenditure_euro / 1000000)

League_simple_money <- ggplot(data = League_Combined_expenditure, aes(x=expenditure_millions, y=Points)) +
  geom_point() + labs(title = "Relationship between Expenditure vs Points Earned",
                      y = "Points Earned",
                      x = "Expenditure (million euros)")
League_simple_money + geom_smooth(method = lm) 

```

##Appendix F

```{r}
coefficients_slr_exp <- lm(Points ~ expenditure_millions, data = League_Combined_expenditure)
coefficients_slr_exp
```

##Appendix G

###Graph 3
```{r}
League_Combined_expenditure2 <- League_Combined_expenditure %>% 
  filter(Expenditure_euro <= 130000000)

League_simple_money2 <- ggplot(data = League_Combined_expenditure2, aes(x=expenditure_millions, y=Points)) +
  geom_point() + labs(title = "Relationship between Expenditure lower than 
                      130 millision euros vs Points Earned",
                      y = "Points Earned",
                      x = "Expenditure (millions of euros)")
League_simple_money2 + geom_smooth(method = lm) 


```

##Appendix H
```{r}
coefficients_slr_exp2 <- lm(Points ~ expenditure_millions, data = League_Combined_expenditure2)
coefficients_slr_exp2

cor(League_Combined_expenditure2$Points, League_Combined_expenditure2$expenditure_millions)
```

##Appendix I
```{r}
#Multiple Linear Regression

League_multiple <- lm(Points ~  Total_Pass + Shots + GA + Poss, data = League_Combined)

summary(League_multiple)


```

##Appendix J
```{r}
League_multiple2 <- lm(Points ~  Total_Pass + Shots + GA , data = League_Combined)
summary(League_multiple2)

coefficients(League_multiple2)
```

##Appendix K 
```{r}
plot(League_multiple2)
```

